The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-08-03 Paper/jaccardbaetica.tex
add new changes xavi

The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
Baetica became an important production and distribution centre during the Roman Empire. However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.

Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD. In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps. Amphoric stamps are used to identify the presence of different groups that might share similar stamps. To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in Baetica province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces. Amphoric production in Baetica province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}.

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/general_map}
\caption{Overview of the sites analyzed in this paper. The color defines the three different regions under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}

@@

This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected Baetica with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.

@@

The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in Baetica province. These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
%\xavi{qué information? El código de letras?}. 
Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. The three letters correspond to a initial forms of a name but they can also appear complete. 
%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}. The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}
\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.
\maria{sí, pero creo que yo no he escrito está frase xD}

@@

The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign. A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself. Baetica became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.

@@

Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
%\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}

@@

Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces. This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border. Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of Britannia extended to \textit{Caledonia}.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/britannia}
\caption{Sites in Britannia where Dressel-20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found in the site. A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
\label{britannia}
\end{figure}

>>>>

The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade in \textit{Baetica} province (currently Andalusia). \textit{Baetica} became an important production and distribution centre during the Roman Empire. However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.

Our case of study has been focused on the amphoric production from 1st to 3rd AD. In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps. Amphoric stamps are used to identify the presence of different groups that might share similar stamps. To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres. They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres. Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance. The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes. Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.

@@

This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces. Amphoric production in \textit{Baetica} province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}).

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/general_map}
\caption{Overview of the sites analyzed in this paper. The color defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}

@@

This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.

@@

The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in \textit{Baetica} province. These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}. The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
%\xavi{qué information? El código de letras?}. 
Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. The three letters correspond to a initial forms of a name but they can also appear complete. 
%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}

@@

Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}. The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}
\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.
\maria{sí, pero creo que yo no he escrito está frase xD}

@@

The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign. A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself. \textit{Baetica} became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.

@@

Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia}).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
%\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}

@@

Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces. This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border. Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of \textit{Britannia} extended to \textit{Caledonia}.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/britannia}
\caption{Sites in \textit{Britannia} where Dressel-20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found in the site. A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
\label{britannia}
\end{figure}
